---
title: Preppin' Data Week 1
author: Scott
date: '2021-06-08'
slug: preppin-data-week-1
categories: []
tags:
  - prepping-data
---

Week 1 requires the ability to read in data, split and clean columns and some date manipulation.

The challenge can be found [here](https://preppindata.blogspot.com/2021/01/2021-week-1.html)


First, load libraries and read in data

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)

dat <- read_excel('week1-data.xlsx')
```


Next, use separate to split store and bike

```{r split-store-bike}
dat_q1 <- separate(dat, `Store - Bike`, into = c('Store', 'Bike'), sep = ' - ') 
  
head(dat_q1)
```

Clean up the `Bike` field with `case_when` from `dplyr`

```{r clean-bike-field}
dat_q2 <- dat_q1 %>% 
  mutate(Bike = case_when(str_starts(Bike, 'Ro') == TRUE ~ 'Road',
                          str_starts(Bike, 'Mount') == TRUE ~ 'Mountain',
                          str_starts(Bike, 'Grav') == TRUE ~ 'Gravel',
                          TRUE ~ 'OTHER'))

head(dat_q2)

```

Create two different date fields, `quarter` and `day of month`. The `lubridate` package makes this really simple.

```{r create-data-fields}
dat_q3 <- dat_q2 %>% 
  mutate(quarter = quarter(Date),
         day_of_month = day(Date))

head(dat_q3)
```

Remove the first 10 orders

```{r remove-first-10}
dat_q4 <- slice(dat_q3, -(1:10))

head(dat_q4)
```

Write data to csv file

```{r write-csv}
data.table::fwrite(dat_q4, file = 'correct-output.csv')
```


This could all be done in one piped call if intended

```{r one-pipe, eval=FALSE}
dat %>% 
  separate(`Store - Bike`, into = c('Store', 'Bike'), sep = ' - ') %>% 
  mutate(Bike = case_when(str_starts(Bike, 'Ro') == TRUE ~ 'Road',
                          str_starts(Bike, 'Mount') == TRUE ~ 'Mountain',
                          str_starts(Bike, 'Grav') == TRUE ~ 'Gravel',
                          TRUE ~ 'OTHER'),
         quarter = quarter(Date),
         day_of_month = day(Date)) %>% 
  slice(-(1:10)) %>% 
  data.table::fwrite(file = 'week1-correct-output.csv')
  
   
```
