---
title: Preppin' Data Week 2
author: Scott
date: 2021-06-05
slug: preppin-data-week-2
categories:
  - R
tags:
  - prepping-data
---
For week 2 there are lots of calculations to be done which can be done with `dplyr` and the `mutate`, `group_by` and `summarise` functions.

The challenge can be found [here](https://preppindata.blogspot.com/2021/01/2021-week-2.html)

First load libraries and read in data.

```{r load-libraries, message=FALSE,warning=FALSE}
library(tidyverse)
library(readxl)
library(writexl)

dat <- read_excel('week2-data.xlsx')
```

Next, clean up the `model` field to leave only the letters. This is done with the regex [aA-zZ]

```{r clean-model-field}
dat_q1 <- mutate(dat, Brand = str_extract(Model, "[aA-zZ]+"))
```

Work out order value by calculating quantity x value per bike

```{r order-value}
dat_q2 <- mutate(dat_q1, order_value = Quantity * `Value per Bike`)
```

Aggregate quantity sold, order value, average value by brand and bike type 

```{r aggregate-1}
dat_q3 <- dat_q2 %>% 
  group_by(`Bike Type`, Brand) %>% 
  summarise(quantity_sold = sum(Quantity), 
            order_value = sum(order_value),
            average_value = mean(sum(order_value)/sum(Quantity)))
```

Calculate days to ship with shipping and order dates

```{r days-to-ship}
dat_q4 <- dat_q2 %>% 
  mutate(days_to_ship = `Shipping Date`- `Order Date`)
```

Aggregate order value, quantity sold and average days to ship by brand and store

```{r aggregate-2}
dat_q5 <- dat_q4 %>% 
  group_by(Brand, Store) %>% 
  summarise(quantity_sold = sum(Quantity), 
            order_value = sum(order_value),
            average_shipping_time = as.numeric(mean(days_to_ship)))
```


Round to 1 decimal place

```{r round}
dat_q6_1 <- mutate(dat_q5, across(where(is.numeric), round, 1))

dat_q6_2 <- mutate(dat_q3, across(where(is.numeric), round, 1))
```

Output

```{r export}
export_data <- list(dat_q6_1, dat_q6_2)

write_xlsx(export_data, path = 'week2-correct-output.xlsx', format_headers = FALSE)
```

Both of these could be written within one pipe by chaining the statements.

```{r one-pipe}
dat %>% 
  mutate(Brand = str_extract(Model, "[aA-zZ]+"),
         order_value = Quantity * `Value per Bike`) %>% 
  group_by(`Bike Type`, Brand) %>% 
  summarise(quantity_sold = sum(Quantity), 
            order_value = sum(order_value),
            average_value = mean(sum(order_value)/sum(Quantity))) %>% 
  mutate(across(where(is.numeric), round, 1)) %>% 
  head()

dat %>% 
  mutate(Brand = str_extract(Model, "[aA-zZ]+"),
         order_value = Quantity * `Value per Bike`,
         days_to_ship = `Shipping Date`- `Order Date`) %>% 
  group_by(Brand, Store) %>% 
  summarise(quantity_sold = sum(Quantity), 
            order_value = sum(order_value),
            average_shipping_time = as.numeric(mean(days_to_ship))) %>% 
   mutate(across(where(is.numeric), round, 1)) %>% 
  head()
  
```


