---
title: Preppin' Data Week 4
author: 'Scott Gee'
date: '2021-06-07'
slug: preppin-data-week-4
categories: 
  - R
tags:
  - prepping-data
---

Week 4 is similar at the start to week 3 but requires the use of joins for the first time in these challenges. This can be done with the `inner_join` function in `dplyr`. If you know SQL, the syntax will be simple to pick up.

The challenge can be found [here](https://preppindata.blogspot.com/2021/01/2021-week-4.html)

First load libraries and read in data. We are using `map_df` again to read in multiple sheets. If you haven't looked at how to do this in week 3 head [there](/2021/06/08/preppin-data-week-3/) before reading this week.

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(data.table)

sheet_names <- excel_sheets('week4-data.xlsx')[1:5]

sheet_names <- set_names(sheet_names, sheet_names) #this ensures that we can use an id column as the store name

dat <- map_df(sheet_names, ~ read_excel('week4-data.xlsx', sheet = .x), .id = 'store')
```

Next pivot the product columns.
```{r pivot}
dat_q2 <- pivot_longer(dat, `New - Saddles`:`Existing - Bags`)
```

Split field with customer type and product using `separate` and rename `value` column to products sold

```{r split-field}
dat_q3 <- separate(dat_q2, name, into = c('customer_type', 'product'), sep = ' - ')

dat_q3 <- rename(dat_q3, products_sold = value)
```

Turn the `date` into a `quarter` with `lubridate`

```{r date-quarter}
dat_q4 <- mutate(dat_q3, quarter = quarter(Date))
```

Sum up products sold by `store` and `quarter`

```{r sum-products}
dat_q5 <- dat_q4 %>% 
  group_by(store, quarter) %>% 
  summarise(products_sold = sum(products_sold))
```

Add targets data

```{r add-targets}
targets <- read_excel('week4-data.xlsx', sheet = 'Targets')
```

Join `targets` data to aggregated stores data



```{r join, message=FALSE}
dat_q7 <- inner_join(dat_q5, targets, by = c('store' = 'Store', 'quarter' = 'Quarter'))
```

Calculate variance between actual sales and target


```{r calculate-variance, message=FALSE}
dat_q9 <- mutate(dat_q7, variance_to_target = Target - products_sold)
```


Rank the stores based on variance to target in each quarter


```{r rank-stores}
dat_q10 <- dat_q9 %>% 
  group_by(quarter) %>% 
  mutate(rank = rank(desc(variance_to_target)))
```


Export data


```{r export}
fwrite(dat_q10, 'correct-output.csv')
```


You could write this within one pipe call with the below:


```{r one-pipe}
one_pipe <- dat %>% 
  pivot_longer(`New - Saddles`:`Existing - Bags`) %>% 
  separate(name, into = c('customer_type', 'product'), sep = ' - ') %>% 
  rename(products_sold = value) %>% 
  mutate(quarter = quarter(Date)) %>% 
  group_by(store, quarter) %>% 
  summarise(products_sold = sum(products_sold)) %>% 
  inner_join(targets, by = c('store' = 'Store', 'quarter' = 'Quarter')) %>% 
  mutate(variance_to_target = Target - products_sold) %>% 
  group_by(quarter) %>% 
  mutate(rank = rank(desc(variance_to_target))) 
    
```

