---
title: Preppin' Data Week 6
author: Scott
date: '2021-06-08'
slug: preppin-data-week-6
categories:
  - R
tags:
  - prepping-data
---
For week 6 it showed how little code can be used to manipulate and summarise data in R using `dplyr`. `pivot_longer` and `pivot_wider` were also used to restructure data in to the required format.

First load libraries and read in data. Removing the dollar sign with `str_remove_all`. The `^` will remove everything but digits (0-9).


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(data.table)

dat <- read_excel('week6-data.xlsx')

#remove dollar sign
dat <- dat %>% 
  mutate(MONEY = str_remove_all(MONEY, "[^0-9]"),
         MONEY = as.integer(MONEY))
```

Create a rank by money earned on all tours, then `group_by` the tour and create a tour rank. After this the summary statistics can be created with calculations. 

`n() will give the count, if the data is grouped it will count the number of observations within the group(s)`

```{r total-tour-money}
dat_q1 <- dat %>% 
  mutate(overall_rank = rank(desc(MONEY))) %>% 
  group_by(TOUR) %>% 
  mutate(tour_rank = rank(desc(MONEY))) %>%  
  summarise(total_prize_money = sum(MONEY),
            total_player = n(),
            total_events = sum(EVENTS),
            average_prize_money = sum(MONEY)/sum(EVENTS),
            average_ranking = mean(overall_rank - tour_rank)) 
  
```

Pivot the data to make it long and then switch the tour names from rows to columns

```{r}
dat_q3 <- dat_q1 %>% 
  pivot_longer(-TOUR) %>% 
  pivot_wider(names_from = TOUR, values_from = value) 
```

The difference between tours can be created with a simple `mutate` call

```{r}
dat_q4 <- dat_q3 %>% 
  mutate(difference = LPGA - PGA) %>% 
  rename(Measure = name)
```

Output the data

```{r}
fwrite(dat_q4, 'week6-correct-output.csv')
```


You can answer this within one chunk of code using multiple pipes like below.

```{r}
dat %>% 
  mutate(overall_rank = rank(desc(MONEY))) %>% 
  group_by(TOUR) %>% 
  mutate(tour_rank = rank(desc(MONEY))) %>%  
  summarise(total_prize_money = sum(MONEY),
            total_player = n(),
            total_events = sum(EVENTS),
            average_prize_money = sum(MONEY)/sum(EVENTS),
            average_ranking = mean(overall_rank - tour_rank)) %>% 
  pivot_longer(-TOUR) %>% 
  pivot_wider(names_from = TOUR, values_from = value) %>% 
  mutate(difference = LPGA - PGA) %>% 
  rename(Measure = name)
```

