---
title: Preppin' Data Week 3
author: Scott
date: '2021-06-08'
slug: preppin-data-week-3
categories:
  - R
tags:
  - prepping-data
---

Week 3 is mostly calculations that can be done with `dplyr`. 

The challenge can be found [here](https://preppindata.blogspot.com/2021/01/2021-week-3.html)

First load libraries and read in data. We can use `map_df` from the `purrr` package to read in multiple sheets from the same excel file. To use you will need a list, in which the objects you want to be evaluated in a function one-by-one. 

Below we create a list of the names of the sheets in the excel file and then use the `read_excel` function with each sheet defined as `.x`. To make this a dataframe we use `map_df` and create a store column with the names of the list that have been created with `set_names`

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(data.table)

sheet_names <- excel_sheets('week3-data.xlsx')

sheet_names <- set_names(sheet_names, sheet_names) #this ensures that we can use an id column as the store name

dat <- map_df(sheet_names, ~ read_excel('week3-data.xlsx', sheet = .x), .id = 'store')
```

Next pivot new and existing and split into customer type and product


```{r pivot-new-existing}
dat_q1 <- dat %>% 
  pivot_longer(`New - Saddles`:`Existing - Bags`, names_to = 'product_type', values_to = 'sales') %>% 
  separate(product_type, into = c('customer_type', 'product')) 
  
```

Rename `sales` to `Products Sold`

```{r rename-column}
dat_q2 <- rename(dat_q1, `Products Sold` = sales)

```

```{r}
dat_q5 <- mutate(dat_q2, quarter = quarter(Date))
```

Get aggregated products sold by:
`Product`, `Quarter`

```{r aggregate-1}
dat_q6_1 <- dat_q5 %>% 
  group_by(product, quarter) %>% 
  summarise(`Products Sold` = sum(`Products Sold`))

```

And then `Store`, `Customer Type`, `Product`


```{r aggregate-2}
dat_q6_2 <- dat_q5 %>% 
  group_by(store, customer_type, product) %>% 
  summarise(`Products Sold` = sum(`Products Sold`))
```


Export data 


```{r}
fwrite(dat_q6_1, file = 'week3-correct-output-1.csv')

fwrite(dat_q6_2, file = 'correct-output-2.csv')
```


These could both be written within one pipe like the below


```{r one-pipe}
dat %>% 
  pivot_longer(`New - Saddles`:`Existing - Bags`, names_to = 'product_type', values_to = 'sales') %>% 
  separate(product_type, into = c('customer_type', 'product')) %>%
  rename(`Products Sold` = sales) %>% 
  mutate(quarter = quarter(Date)) %>% 
  group_by(product, quarter) %>% 
  summarise(`Products Sold` = sum(`Products Sold`)) 

dat %>% 
  pivot_longer(`New - Saddles`:`Existing - Bags`, names_to = 'product_type', values_to = 'sales') %>% 
  separate(product_type, into = c('customer_type', 'product')) %>%
  rename(`Products Sold` = sales) %>% 
  mutate(quarter = quarter(Date)) %>%
  group_by(store, customer_type, product) %>% 
  summarise(`Products Sold` = sum(`Products Sold`))
```
